Time delay estimation in unresolved lensed quasars

### 2020-12-25 paper/aa-package/paper.tex
intro

Time-delays between light curves of Gravitationally Lensed Quasars (GLQs) can be estimated with various methods \cite{Tewes2013}, such as free-knot spline interpolation \cite{free} or via Gaussian Process (GP) regression \cite{Rasmussen}. Such methods work under the assumption that multiple images of the same quasar are fully resolved so that a light curve can be extracted for each of them. However, most known GLQs \cite{quasars_db} have small angular separations $\Delta\theta < \sim{3}{arcsec}$. 
This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent.
%This would make big telescopes, with high angular resolution, the ideal instruments to perform lensed quasars monitoring. Furthermore, these telescopes are placed in geographical areas where effects of atmospheric turbulence are less prominent. 
However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years \cite{Cosmograil2020}. Therefore, due to the amount of observational requests that big telescopes have to fulfill, they can not be employed for these purposes. On the other hand, small/medium sized telescopes ($1$-$2$m) \cite{Borgeest1996} can be used. Unfortunately their reduced angular resolution, together with their often less privileged geographical positions in terms of atmospheric seeing, may worsen the effective angular resolution up to $3 arcsec$ \cite{karttunen2016fundamental}. In addition, small telescopes cannot observe faint distant sources. These shortcomings prevent the acquisition of a large number of GLQ datasets suitable for the application of the aforementioned techniques \cite{free, Rasmussen}.

>>>>

In this respect, the fractional error of $H_0$, for an ensemble of N Gravitationally Lensed Quasars (GLQs), is related to the uncertainties in the time-delay estimation $\sigma_{\Delta T}$, line-of-sight convergence $\sigma_{los}$ and lens surface density $\sigma_{\braket{k}}$ as [\cite{Tie_2017}]:
\frac{\sigma_H^2}{H_0^2} \sim \frac{\sigma_{\Delta T}^2/\Delta T^2 + \sigma_{los}^2 + \sigma_{\braket{k}}^2}{N}
where the first two terms are dominated by random errors and their contribution to the overall error scales as $N^{-1/2}$. Therefore, increasing the sample of analysed GLQs allows to reduce the error on $H_0$. 
In this paper we focus on the time delay estimation and its relative error. To date, a sample of about $220$ GLQs is available\footnote{https://research.ast.cam.ac.uk/lensedquasars/index.html}. However, only a very small subset, consisting of the brightests and with with well separated multiple images, has been used to measure $H_0$. The reason relies on the fact that it is easier and safer to extract information, and consequently to reduce the error on $H_0$, for such systems.
In this respect, Fig. \ref{fig:qsosepmagboth} shows the distribution of the multiple images versus the maximum image separation for the known GLQs. Systems being part of the grey region, which represent a consistent fraction of the total sample, can be regarded as not optimal candidates, since they are less bright and/or have very small image separations. 
\begin{figure}[h]
	\includegraphics[width=1.0\linewidth]{Figures/QSO_sep_mag_both}
	\caption[cm]{comments}
	\label{fig:qsosepmagboth}
In addition, [\cite{shu2020discovering}] have presented a new method to find GLQ systems from unresolved light curves. Such systems have therefore very small-separation between multiple images. This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent. However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years [\cite{Cosmograil2020}]. Therefore, due to the amount of observational requests that big telescopes have to fulfil, they can not be employed for these purposes. On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used. Unfortunately, their reduced angular resolution is further worsened by their often less privileged geographical positions in terms of atmospheric seeing, up to 3 arcsec \cite{karttunen2016fundamental}. In addition, small telescopes cannot observe faint distant sources. 
